---
title: "WorkSheet2"
author: "Neil Francis N. Navarro"
date: "2022-10-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

**Worksheet for R Programming**

>**Instructions:**

+ 
> Use RStudio or the RStudio Cloud accomplish this worksheet. + Save the R script as RWorksheet_lastname#2.R.

+ 
>Create your own GitHub repository and push the R script as well as this pdf worksheet
to your own repo.

>Accomplish this worksheet by answering the questions being asked and writing the code
manually

+
>**Using Vectors**
>1. Create a vector using : operator
>a. Sequence from -5 to 5. Write the R code and its output. Describe its output.

```{r x1A}
seq(-5:5)
```
>b. x <- 1:7. What will be the value of x?

```{r x1B}
x <- 1:7
x
```
>2.* Create a vector using seq() function
>a. seq(1, 3, by=0.2) # specify step size
Write the R code and its output. Describe the output.

```{r x2}
seq(1, 3, by=0.2)
```
>The result will shows numbers that is being added by 0.2 from 1 to 3.


>3. A factory has a census of its workers. There are 50 workers in total. The following
list shows their ages: 34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18.

```{r x3}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37,
            34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 
            35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 
            61, 54, 58, 26,18)
ages
```

>a. Access 3rd element, what is the value?

```{r x3A}
ages[3]
```
>b. Access 2nd and 4th element, what are the values?

```{r x3B}
ages[2]
ages[4]
```
>c. Access all but the 1st element is not included. Write the R code and its output.

```{r x3C}
ages[2:50]
```

>4. *Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the vector,
names(x).

```{r x4}
names <- c("first"=3, "second"=0, "third"=9)
names
```
>a. Print the results. Then access x[c("first", "third")].
>Describe the output.

```{r x4A}
names[c("first", "third")]
```
>b. Write the code and its output.

```{r x4B}
names[c("first", "third")]
```

>5. Create a sequence x from -3:2.

```{r x5}
x <- c(-3:2)
x
```
>a. Modify 2nd element and change it to 0;
>x[2] <- 0
>x
>Describe the output.

```{r x5A}
x <- 0
x
```
>b. Write the code and its output.

```{r x5B}
x <- 0
x
```

6. *The following data shows the diesel fuel purchased by Mr. Cruz.

```{r 6}
deisel <- data.frame(
  
  Months = c("Jan", "Feb", "March", "Apr", "May", "June"),
  Php = c(52.50, 57.25, 60.00, 65.00,74.25, 54.00),
  Liter = c(25, 30, 40, 50, 10, 45),
  stringsAsFactors = FALSE
                    )
deisel
```
a. Create a data frame for month, price per liter (php) and purchase-quantity (liter).
Write the codes.

```{r 6A}
deisel$Php
deisel$Liter
```
>b. What is the average fuel expenditure of Mr. Cruz from Jan to June? Note: Use
weighted.mean(Liter, Php)

>weighted.mean(Liter,Php)

>[**1] 32.65152**

```{r 6B}
deisel <- data.frame(
  
  Months <- c("Jan", "Feb", "March", "Apr", "May", "June"),
  Php <- c(52.50, 57.25, 60.00, 65.00,74.25, 54.00),
  Liter <- c(25, 30, 40, 50, 10, 45),
  stringsAsFactors = FALSE
                    )
weighted.mean(Liter,Php)
```
>7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths
(in miles) of 141 “major” rivers in North America, as compiled by the US Geological
Survey”.

```{r 7}

```
>a. Type “rivers” in your R console. Create a vector data with 7
elements, containing the number of elements (length) in rivers,
their sum (sum), mean (mean), median (median), variance (var)
standard deviation (sd), minimum (min) and maximum (max).
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers),
sd(rivers), min(rivers), max(rivers))

```{r 7A}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), 
          var(rivers),sd(rivers), min(rivers), max(rivers))
data
```
>b. What are the results?

```{r 7B}
data
```
>c. Write the code and its outputs.


```{r 7C}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), 
          var(rivers),sd(rivers), min(rivers), max(rivers))
data
```
>8. The table below gives the 25 most powerful celebrities and their annual pay as ranked
by the editions of Forbes magazine and as listed on the Forbes.com website.
Figure 1: Forbes Ranking

```{r 8}
 PowerRanking <- 1:25
  Celebrities <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods",
          "Steven Spielherg", "Howard Stern", "50 Cent", "Cast of the Sopranos",
          "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali",
          "Paul McCartney","George Lucas","Elton John","David Letterman","Phil Mickelson",
          "J.K Rowling","Bradd Pitt","Peter Jackson","Dr. Phil McGraw",
          "Jay Lenon","Celine Dion","Kobe Bryant")
  Pay <- c(67,90,225,110,90,332,302,41,52,88,55,44,55,40,233,34,40,47,75,25,39,45,32,40,31)
  
  Ranking <- data.frame(PowerRanking, Celebrities, Pay)
  Ranking
```
>a. Create vectors according to the above table. Write the codes.

```{r 8A}
 PowerRanking <- 1:25
  Celebrities <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods",
          "Steven Spielherg", "Howard Stern", "50 Cent", "Cast of the Sopranos",
          "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali",
          "Paul McCartney","George Lucas","Elton John","David Letterman","Phil Mickelson",
          "J.K Rowling","Bradd Pitt","Peter Jackson","Dr. Phil McGraw",
          "Jay Lenon","Celine Dion","Kobe Bryant")
  Pay <- c(67,90,225,110,90,332,302,41,52,88,55,44,55,40,233,34,40,47,75,25,39,45,32,40,31)
  
  Ranking <- data.frame(PowerRanking, Celebrities, Pay)
  Ranking
```
>b. Modify the power ranking and pay of J.K. Rowling. Change power ranking to 15 and
pay to 90. Write the codes and its output.

```{r 8B}
PowerRanking[19] <- 15;
  PowerRanking  
  
  Pay[19] <- 90;
  Pay

  ArtistRanking <- data.frame(PowerRanking, Celebrities, Pay)
  ArtistRanking  
```
>c. Interpret the data.

```{r 8C}
PowerRanking[19] <- 15;
  PowerRanking  
  
  Pay[19] <- 90;
  Pay

  ArtistRanking <- data.frame(PowerRanking, Celebrities, Pay)
  ArtistRanking  
```





